\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the
%first footnote. If that is unneeded, please comment it out.

%%% PACKAGES
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts,bm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{amsthm}
\usepackage{coqdoc}


%%% COMMMANDS
\newcommand{\macheps}{\epsilon_{\mathrm{mach}}}
\newcommand{\tmacheps}{\tilde{\epsilon}_{\mathrm{mach}}}
\newcommand{\unlambda}{\lambda_{\mathrm{un}}}
\newcommand{\etamax}{\eta_{\mathrm{max}}}
\newcommand{\fl}{\operatorname{fl}}
\newcommand{\vc}[1]{\textbf{#1}}


\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{prop}[theorem]{Proposition}
\newtheorem{cor}{Corollary}
\newtheorem{defn}{Definition}[section]
\newtheorem{conj}{Conjecture}[section]
\newtheorem{exmp}{Example}[section]
\newtheorem{rem}{Remark}

%%% CODE
\usepackage{listings}
\usepackage{lstlangcoq}
\usepackage{xcolor}
\usepackage[T1]{fontenc}
\usepackage[scaled]{beramono}



\lstset{mathescape=true,language=Coq}

    
%%% DOCUMENT  
\begin{document}
\title{LAProof: A Library of Formal Proofs of Accuracy and
  Correctness for Linear Algebra Programs
  % .\\
% {\footnotesize \textsuperscript{*}Note: Sub-titles are not
% captured in Xplore and should not be used}
% \thanks{Identify applicable funding agency here. If none, delete this.}
}

%\author{Authors omitted for double-blind review}

\newcommand\comment[1]{}

\author{
\IEEEauthorblockN{
Ariel E. Kellison\IEEEauthorrefmark{1}\IEEEauthorrefmark{4},
Andrew W. Appel\IEEEauthorrefmark{2},
Mohit Tekriwal\IEEEauthorrefmark{3},
David Bindel\IEEEauthorrefmark{1}}
    \IEEEauthorblockA{\IEEEauthorrefmark{1}
Dept. of Computer Science, Cornell University, Ithaca, NY, USA }
    \IEEEauthorblockA{\IEEEauthorrefmark{3}
Dept. of Aerospace Engineering, University of Michigan, Ann Arbor, MI, USA}
    \IEEEauthorblockA{\IEEEauthorrefmark{2} 
Dept. of Computer Science, Princeton University, Princeton, NJ, USA }
    \IEEEauthorblockA{\IEEEauthorrefmark{4}
 Sandia National Laboratories, Livermore, CA, USA }
}


\maketitle
\thispagestyle{plain}\pagestyle{plain} % turn on page numbers for submission

\begin{abstract}
The LAProof library provides formal machine-
checked proofs of the accuracy of basic linear algebra operations:
inner product using conventional multiply and add, inner product
using fused multiply-add, scaled matrix-vector and matrix-matrix
multiplication, and scaled vector and matrix addition. These
proofs can connect to concrete implementations of low-level basic
linear algebra subprograms; as a proof of concept we present a
machine-checked correctness proof of a C function implementing
sparse matrix-vector multiplication using the compressed sparse
row format. Our accuracy proofs are backward error bounds
and mixed backward-forward error bounds that account for
underflow, proved subject to no assumptions except a low-level
formal model of IEEE-754 arithmetic. We treat low-order error
terms concretely, not approximating as $\mathcal{O}(u^2)$.
\end{abstract}

\begin{IEEEkeywords}
rounding error analysis, formal verification,
floating-point arithmetic, program verification
\end{IEEEkeywords}

\input{introduction}

\input{lib}

\input{models}

\input{sparse}

\input{relatedwork}

\input{conclusion}



 \section{Acknowledgments}  
This research was supported in part by NSF Grants CCF-
2219997 and CCF-2219757, by a U.S. Department of Energy
Computational Science Fellowship DE-SC0021110, and by the
Chateaubriand fellowship program.

Sandia National Laboratories is a multimission laboratory
managed and operated by National Technology \& Engineering
Solutions of Sandia, LLC, a wholly owned subsidiary of Hon-
eywell International Inc., for the U.S. Department of Energy’s
National Nuclear Security Administration under contract DE-
NA0003525.


\bibliographystyle{IEEEtran}
\bibliography{bib.bib}


\end{document}
